---
title: "Eb_LLBehavior_Strikes"
output: html_document
---


```{r}
#Analysis for strike order and number of un/sucessful strikes 

library(lattice)
library(readxl)

#Loaded Sheet
pstrikes = read_excel("Eb_Strikes_Platforms.xlsx")
summary(pstrikes) 
```

```{r}
##Subsetting the data 
pstrikes20 <- subset(pstrikes, pstrikes$YEAR == "2020")
summary(pstrikes20)

controlstrikes<- subset(pstrikes20, pstrikes20$MODALITY == "C")
chemstrikes<- subset(pstrikes20, pstrikes20$MODALITY=="NLV")
lcstrikes<- subset(pstrikes20, pstrikes20$MODALITY=="NV")
vcstrikes<- subset(pstrikes20, pstrikes20$MODALITY=="NL")
```

```{r}
#Normality Testing 
shapiro.test(pstrikes20$SSTRIKES) #not normal
shapiro.test(pstrikes20$FSTRIKES)#not normal

# Fit the mixed-e_ects model for successful strikes 
library(lme4)
model <- lmer(avg_SSTRIKES ~ TREATMENT * LIGHT + (1 | FISH), data = SStrikesAvg)
# Check for singular fit
if (isSingular(model, tol = 1e-4)) {
print("The model is singular.")

} else {
print("The model is not singular.")
}
# Display the summary of the model
summary(model)

# Fit the mixed-e_ects model for failed strikes 
library(lme4)
model1 <- lmer(avg_FSTRIKES ~ TREATMENT * LIGHT + (1 | FISH), data = FStrikesAvg)
# Check for singular fit
if (isSingular(model, tol = 1e-4)) {
print("The model is singular.")

} else {
print("The model is not singular.")
}
# Display the summary of the model
summary(model1)

#Plots for Figure 8
livesstrikes <- pstrikes20 %>% group_by(FISH, LIGHT, TREATMENT) %>%
   summarize(avg_livestrikes=mean(SSTRIKES_LIVE))

#write.csv(FStrikesAvg, file = "StrikesAvg.csv") #Add your own path like to your desktop or somewhere, end it with the file name you want and the extension .csv

p1<- ggplot(livesstrikes, aes(x = factor(LIGHT, level = c('light', 'dark')), y = avg_livestrikes, shape = TREATMENT)) +
geom_point(position = position_dodge(width = 0.5), size = 3) + # Add dodge to separate
scale_fill_manual(values = c("gray", "gray44"))+
  #ablation groups and increase point size
labs(
#title = "Sucessful Strikes by Light Condition and Ablation Status",
x = "Light Condition",
y = "Number of Strikes",
#color = "Ablation Status",
shape = "Ablation Status"
) +
theme_bw() + # Use a minimal theme
theme(
panel.grid.minor = element_blank(),
  legend.position = "bottom" # Move the legend to the bottom
)

p1 

ggsave("LiveSStrikes.tiff", width = 6, height = 4) #This saves it to active directory


deadsstrikes <- pstrikes20 %>% group_by(FISH, LIGHT, TREATMENT) %>%
   summarize(avg_deadsstrikes=mean(SSTRIKES_DEAD))

#write.csv(FStrikesAvg, file = "StrikesAvg.csv") #Add your own path like to your desktop or somewhere, end it with the file name you want and the extension .csv

p4<- ggplot(deadsstrikes, aes(x = factor(LIGHT, level = c('light', 'dark')), y = avg_deadsstrikes, shape = TREATMENT)) +
geom_point(position = position_dodge(width = 0.5), size = 3) + # Add dodge to separate
scale_fill_manual(values = c("gray", "gray44"))+
  #ablation groups and increase point size
scale_y_continuous(limits=c(0,4), breaks=seq(0, 4, 1))+
  labs(
#title = "Sucessful Strikes by Light Condition and Ablation Status",
x = "Light Condition",
y = "Number of Strikes",
#color = "Ablation Status",
shape = "Ablation Status"
) +
theme_bw() + # Use a minimal theme
theme(
panel.grid.minor = element_blank(),
  legend.position = "bottom" # Move the legend to the bottom
)

p4

ggsave("DeadSStrikes.tiff", width = 6, height = 4) #This saves it to active directory


livefstrikes <- pstrikes20 %>% group_by(FISH, LIGHT, TREATMENT) %>%
   summarize(avg_livefstrikes=mean(FSTRIKES_LIVE))

#write.csv(FStrikesAvg, file = "StrikesAvg.csv") #Add your own path like to your desktop or somewhere, end it with the file name you want and the extension .csv

p2<- ggplot(livefstrikes, aes(x = factor(LIGHT, level = c('light', 'dark')), y = avg_livefstrikes, shape = TREATMENT)) +
geom_point(position = position_dodge(width = 0.5), size = 3) + # Add dodge to separate
#scale_fill_manual(values = c("gray", "gray44"))+
  #ablation groups and increase point size
#ylim(c(0,7))+
  scale_y_continuous(limits=c(0,7), breaks=seq(0, 7, 1))+
  labs(
    #title = "Sucessful Strikes by Light Condition and Ablation Status",
x = "Light Condition",
y = "Number of Strikes",
#color = "Ablation Status",
shape = "Ablation Status"
) +
theme_bw() + # Use a minimal theme
theme(
panel.grid.minor = element_blank(),
  legend.position = "bottom", # Move the legend to the bottom
axis.ticks.length = unit(1, "mm"),
)

p2 

ggsave("LiveFStrikes.tiff", width = 6, height = 4) #This saves it to active directory


deadfstrikes <- pstrikes20 %>% group_by(FISH, LIGHT, TREATMENT) %>%
   summarize(avg_deadfstrikes=mean(FSTRIKES_DEAD))

#write.csv(FStrikesAvg, file = "StrikesAvg.csv") #Add your own path like to your desktop or somewhere, end it with the file name you want and the extension .csv

p3<- ggplot(deadfstrikes, aes(x = factor(LIGHT, level = c('light', 'dark')), y = avg_deadfstrikes, shape = TREATMENT)) +
geom_point(position = position_dodge(width = 0.5), size = 3) + # Add dodge to separate
#scale_fill_manual(values = c("gray", "gray44"))+
  #ablation groups and increase point size
#ylim(c(0,10))+
  scale_y_continuous(limits = c(0, 7), breaks = round(seq(0,7, 1),1))+
  #expand_limits(y = 6)+
  labs(
    #title = "Sucessful Strikes by Light Condition and Ablation Status",
x = "Light Condition",
y = "Number of Strikes",
#color = "Ablation Status",
shape = "Ablation Status"
) +
theme_bw() + # Use a minimal theme
theme(
panel.grid.minor = element_blank(),
  legend.position = "bottom", # Move the legend to the bottom
axis.ticks.length = unit(1, "mm"),
)

p3

ggsave("DeadFStrikes.tiff", width = 6, height = 4) #This saves it to active directory
```



```{r}
##Mean preference score analysis - were live or dead prey detected earlier within each treatment? 
#Averaging the scores from all three light trials and all three dark trials, per request of editor 
##LiveprefAvg <- strikes20 %>% group_by(FISH, LIGHT, TREATMENT) %>%
   summarize(avg_livepref=mean(AVG_LIVE),
             std_livepref=sd(AVG_LIVE))

DeadprefAvg <- strikes20 %>% group_by(FISH, LIGHT, TREATMENT) %>%
   summarize(avg_deadpref=mean(AVG_DEAD),
             std_deadpref=sd(AVG_DEAD))

##Plots used in Figure 8
#Dead Prey 
p5<- ggplot(DeadprefAvg, aes(x = factor(LIGHT, level = c('light', 'dark')), y = avg_deadpref, shape = TREATMENT)) +
geom_point(position = position_dodge(width = 0.5), size = 3) + # Add dodge to separate
#scale_fill_manual(values = c("gray", "gray44"))+
  #ablation groups and increase point size
#ylim(c(0,7))+
  scale_y_continuous(limits=c(0,8), breaks=seq(0, 8, 1))+
  labs(
    #title = "Sucessful Strikes by Light Condition and Ablation Status",
x = "Light Condition",
y = "Preference Score",
#color = "Ablation Status",
shape = "Ablation Status"
) +
theme_bw() + # Use a minimal theme
theme(
panel.grid.minor = element_blank(),
  legend.position = "bottom", # Move the legend to the bottom
axis.ticks.length = unit(1, "mm"),
)

p5

ggsave("DeadPrefAvg.tiff", width = 6, height = 4) #This saves it to active directory

#Live prey 
p6<- ggplot(LiveprefAvg, aes(x = factor(LIGHT, level = c('light', 'dark')), y = avg_livepref, shape = TREATMENT)) +
geom_point(position = position_dodge(width = 0.5), size = 3) + # Add dodge to separate
#scale_fill_manual(values = c("gray", "gray44"))+
  #ablation groups and increase point size
#ylim(c(0,7))+
  scale_y_continuous(limits=c(0,8), breaks=seq(0, 8, 1))+
  labs(
    #title = "Sucessful Strikes by Light Condition and Ablation Status",
x = "Light Condition",
y = "Preference Score",
#color = "Ablation Status",
shape = "Ablation Status"
) +
theme_bw() + # Use a minimal theme
theme(
panel.grid.minor = element_blank(),
  legend.position = "bottom", # Move the legend to the bottom
axis.ticks.length = unit(1, "mm"),
)

p6

ggsave("LivePrefAvg.tiff", width = 6, height = 4) #This saves it to active directory

##
# Fit the mixed-effects model for live prey preference scores 
library(lme4)
library(lmerTest)
model3 <- lmer(avg_livepref ~ TREATMENT * LIGHT + (1 | FISH), data = LiveprefAvg)
# Check for singular fit
if (isSingular(model, tol = 1e-4)) {
print("The model is singular.")

} else {
print("The model is not singular.")
}
# Display the summary of the model
summary(model3)

#Model for dead prey preference scores 
library(lme4)
library(lmerTest)
model4 <- lmer(avg_deadpref ~ TREATMENT * LIGHT + (1 | FISH), data = DeadprefAvg)
# Check for singular fit
if (isSingular(model, tol = 1e-4)) {
print("The model is singular.")

} else {
print("The model is not singular.")
}
# Display the summary of the model
summary(model4)



```

```{r}
##Did the platform's time in tank affect the preference for live or dead prey? 

library(lattice)
library(readxl)

library(tidyverse)
library(rstatix)

#Loaded Sheet 
platforms = read_excel("Eb_Platforms.xlsx")
summary(platforms)

#Subsetting the data 
platforms20 <- subset(platforms, platforms$YEAR == "2020")

##Was there an overall affect of platform time on preference for live prey 
kruskal.test(AVG_LIVESCORE ~ ALL_PLATFORMS_TANKTIME, data=platforms20) ##NS, p-value = 0.5586

##Was there an overall affect of platform time on preference for dead prey 
kruskal.test(AVG_DEADSCORE ~ ALL_PLATFORMS_TANKTIME, data=platforms20) ##NS, p-value = 0.3666


##PLATFORM 1
shapiro.test(platforms20$P1_LIVE)

kruskal.test(P1_LIVE ~ P1_TANKTIME, data = platforms20) ##Not sig, p-value = 0.4042

a14<- aov(data = platforms20, P1_LIVE ~ as.factor(P1_TANKTIME) * as.factor(MODALITY))
summary(a14) #NS


#PLATFORM 2
shapiro.test(platforms20$P2_LIVE)

kruskal.test(P2_LIVE ~ P2_TANKTIME, data = platforms20) ##p-value = 0.2182

a15<- aov(data = platforms20, P2_LIVE ~ as.factor(P2_TANKTIME) * as.factor(MODALITY))
summary(a15) #NS


##PLATFORM 3
shapiro.test(platforms20$P3_LIVE)

kruskal.test(P3_LIVE ~ P3_TANKTIME, data = platforms20) ##p-value = 0.5866

a16<- aov(data = platforms20, P3_LIVE ~ as.factor(P3_TANKTIME) * as.factor(MODALITY))
summary(a16) #Only interaction significant 

ph7 <- glht(a16, linfct = mcp("as.factor(MODALITY)" = "Tukey"))
summary(ph7) ##NS after correction 

ph8 <- glht(a16, linfct = mcp("as.factor(P3_TANKTIME)" = "Tukey"))
summary(ph8) ##Only 2T vs. 1T is sig., but 1T vs. 3T isn't
#Agree w/ AM that since 3T vs. 1T isn't significant this isn't convincing 


##PLATFORM 4
shapiro.test(platforms20$P4_LIVE)

kruskal.test(P4_LIVE ~ P4_TANKTIME, data = platforms20) ##p-value = 0.1464

a17<- aov(data = platforms20, P4_LIVE ~ as.factor(P4_TANKTIME) * as.factor(MODALITY))
summary(a17) #NS

```
##Old strike analysis based on nonparametric data, editor required it not be used 
##Testing whether strikes on live vs. dead prey varied with modality 
kruskal.test(SSTRIKES ~ MODALITY, data = pstrikes20) #p-value = 0.1399

kruskal.test(SSTRIKES_LIVE ~ MODALITY, data = pstrikes20) #p-value = 0.09755

kruskal.test(SSTRIKES_DEAD ~ MODALITY, data = pstrikes20) #p-value = 0.07765

kruskal.test(FSTRIKES ~ MODALITY, data = pstrikes20) # p-value = 0.174

kruskal.test(FSTRIKES_DEAD ~ MODALITY, data = pstrikes20) #p-value = 0.09376

kruskal.test(FSTRIKES_LIVE ~ MODALITY, data = pstrikes20) #p-value = 0.5966
```

##Old plots for Fig. 9
```{r}
##Plots for Figure 9

library(ggplot2)
library(tidyverse)

strikes_new <- pstrikes20    # Duplicate data
modality <- factor(strikes_new$MODALITY, c("C", "NL", "NV", "NLV"))

plot13 <- ggplot(data=pstrikes20, aes(x=modality, y=SSTRIKES_LIVE)) + #geom_boxplot(outlier.shape = 1) +
geom_jitter(color="black", shape=2, size=2, width = 0.2, height = 0) +      
stat_summary(fun.y=mean, colour="black", geom="point", 
                           shape=20, size=3,show_guide = FALSE)+
    #scale_color_manual(values = c("darkorchid2", "darkturquoise", "darkorchid4", "deepskyblue")) +
  #scale_y_continuous(breaks = round(seq(100,2000, by = 500),1))+
  xlab("Modality") +
  ylab("# of Sucessful Strikes") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_line(color="black"),
        axis.ticks.length = unit(1, "mm"),
        axis.text.x = element_text(color="black", size = 11, angle = 0, vjust = 1, hjust = 0.5),
        axis.text.y = element_text(color="black", size = 11),
        legend.position="none")
  
plot13 


plot14 <- ggplot(data=pstrikes20, aes(x=modality, y=SSTRIKES_DEAD)) + #geom_boxplot(outlier.shape = 1) +
geom_jitter(color="black", shape=2, size=2, width = 0.2, height = 0) +          
stat_summary(fun.y=mean, colour="black", geom="point", 
                           shape=20, size=3,show_guide = FALSE)+
    #scale_color_manual(values = c("darkorchid2", "darkturquoise", "darkorchid4", "deepskyblue")) +
  #scale_y_continuous(breaks = round(seq(100,2000, by = 500),1))+
  xlab("Modality") +
  ylab("# of Sucessful Strikes") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_line(color="black"),
        axis.ticks.length = unit(1, "mm"),
        axis.text.x = element_text(color="black", size = 11, angle = 0, vjust = 1, hjust = 0.5),
        axis.text.y = element_text(color="black", size = 11),
        legend.position="none")
  
plot14 

plot15 <- ggplot(data=pstrikes20, aes(x=modality, y=FSTRIKES_DEAD)) + #geom_boxplot(outlier.shape = 1) +
geom_jitter(color="black", shape=2, size=2, width = 0.2, height = 0) +          
stat_summary(fun.y=mean, colour="black", geom="point", 
                           shape=20, size=3,show_guide = FALSE)+
    #scale_color_manual(values = c("darkorchid2", "darkturquoise", "darkorchid4", "deepskyblue")) +
  #scale_y_continuous(breaks = round(seq(100,2000, by = 500),1))+
  xlab("Modality") +
  ylab("# of Failed Strikes") +
  ylim(0, 11) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_line(color="black"),
        axis.ticks.length = unit(1, "mm"),
        axis.text.x = element_text(color="black", size = 11, angle = 0, vjust = 1, hjust = 0.5),
        axis.text.y = element_text(color="black", size = 11),
        legend.position="none")
  
plot15 

plot16 <- ggplot(data=pstrikes20, aes(x=modality, y=FSTRIKES_LIVE)) + #geom_boxplot(outlier.shape = 1) +
geom_jitter(color="black", shape=2, size=2, width = 0.2, height = 0) +       
stat_summary(fun.y=mean, colour="black", geom="point", 
                           shape=20, size=3,show_guide = FALSE)+
    #scale_color_manual(values = c("darkorchid2", "darkturquoise", "darkorchid4", "deepskyblue")) +
  #scale_y_continuous(breaks = round(seq(100,2000, by = 500),1))+
  xlab("Modality") +
  ylab("# of Failed Strikes") +
   ylim(0, 11) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_line(color="black"),
        axis.ticks.length = unit(1, "mm"),
        axis.text.x = element_text(color="black", size = 11, angle = 0, vjust = 1, hjust = 0.5),
        axis.text.y = element_text(color="black", size = 11),
        legend.position="none")
  
plot16 


#Previous analysis of preference scores based on nonparametric data, editor required it to not be used 
wilcox.test(controlstrikes$AVG_DEAD, controlstrikes$AVG_LIVE, paired = T) ## p-value = 0.2249

wilcox.test(chemstrikes$AVG_DEAD, chemstrikes$AVG_LIVE, paired = T) ##p-value = 0.8124

wilcox.test(lcstrikes$AVG_DEAD, lcstrikes$AVG_LIVE, paired = T) ##p-value = 0.9687

wilcox.test(vcstrikes$AVG_DEAD, vcstrikes$AVG_LIVE, paired = T) ##p-value = 0.06204

##Did the mean preference score vary among treatments? 
kruskal.test(AVG_LIVE~MODALITY, data = pstrikes20)

kruskal.test(AVG_DEAD~MODALITY, data = pstrikes20)
dunn_test(data = pstrikes20, AVG_DEAD ~ MODALITY, p.adjust.method = "bonferroni")
