---
title: "Eb_LLBehavior_PreyConsumption"
output: html_document
---


```{r}
library(lattice)
library(readxl)

#Loaded Sheet 
prey = read_excel("Eb_Prey_Consumption.xlsx")
summary(prey) 

###Code for modality: C = light trial, LL intact; NV = dark trial, LL intact, NL = light trial, LL ablated, NVL = dark trial, LL ablated
```

```{r}
#Subsetting data 
fish20 <- subset(prey, prey$YEAR == "2020")
summary(fish20)

controlprey <- subset(fish20, fish20$MODALITY == "C")
lcprey <- subset(fish20, fish20$MODALITY == "NV")
chemprey <- subset(fish20, fish20$MODALITY == "NLV")
vcprey <- subset(fish20, fish20$MODALITY == "NL")
```

```{r}
##Testing for normality 
shapiro.test(fish20$TOTAL_PREY)
shapiro.test(fish20$TOTAL_LIVE)
shapiro.test(fish20$TOTAL_DEAD)

#Did number of prey consumed vary with modality? 
kruskal.test(data=fish20, TOTAL_PREY~MODALITY)

kruskal.test(data=fish20, TOTAL_LIVE~MODALITY)
dunn_test(data = fish20, TOTAL_LIVE ~ MODALITY, p.adjust.method = "bonferroni")

kruskal.test(data=fish20, TOTAL_DEAD~MODALITY)
dunn_test(data = fish20, TOTAL_DEAD ~ MODALITY, p.adjust.method = "bonferroni")

#Was the total number of live & dead prey consumed different across all treatments? 
wilcox.test(fish20$TOTAL_LIVE, fish20$TOTAL_DEAD, paired = T) #Yes 
mean(fish20$TOTAL_DEAD)
mean(fish20$TOTAL_LIVE)


#Is # live prey vs. dead prey different for each modality

wilcox.test(controlprey$TOTAL_LIVE, controlprey$TOTAL_DEAD, paired = T) #.01298

wilcox.test(lcprey$TOTAL_LIVE, lcprey$TOTAL_DEAD, paired = T) #p-value = 0.04472

wilcox.test(chemprey$TOTAL_LIVE, chemprey$TOTAL_DEAD, paired = T) ##p-value = 0.3447 NOT SIG INTERESTING

wilcox.test(vcprey$TOTAL_LIVE, vcprey$TOTAL_DEAD, paired = T) # p-value = 0.02627
```



```{r}
##Plots for Figure 10 
```{r}
library(ggplot2)
library(tidyverse)
prey_new <- fish20    # Duplicate data
modality <- factor(prey_new$MODALITY, c("C", "NL", "NV", "NLV"))

plot13 <- ggplot(data=fish20, aes(x=modality, y=TOTAL_DEAD)) + #geom_boxplot(outlier.shape = 1) +
     geom_jitter(color="black", shape=2, size=2, width = 0.2, height = 0) + 
     stat_summary(fun.y=mean, colour="black", geom="point", 
                           shape=20, size=3,show_guide = FALSE)+
    #scale_color_manual(values = c("darkorchid2", "darkturquoise", "darkorchid4", "deepskyblue")) +
  #scale_y_continuous(breaks = round(seq(100,2000, by = 500),1))+
  xlab("Modality") +
  ylab("# Dead Prey Consumed") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_line(color="black"),
        axis.ticks.length = unit(1, "mm"),
        axis.text.x = element_text(color="black", size = 11, angle = 0, vjust = 1, hjust = 0.5),
        axis.text.y = element_text(color="black", size = 11),
        legend.position="none")
  
plot13 

plot14 <- ggplot(data=fish20, aes(x=modality, y=TOTAL_LIVE)) + #geom_boxplot(outlier.shape = 1) +
     geom_jitter(color="black", shape=2, size=2, width = 0.2, height = 0) + 
     stat_summary(fun.y=mean, colour="black", geom="point", 
                           shape=20, size=3,show_guide = FALSE)+
    #scale_color_manual(values = c("darkorchid2", "darkturquoise", "darkorchid4", "deepskyblue")) +
  #scale_y_continuous(breaks = round(seq(100,2000, by = 500),1))+
  xlab("Modality") +
  ylab("# Live Prey Consumed") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_line(color="black"),
        axis.ticks.length = unit(1, "mm"),
        axis.text.x = element_text(color="black", size = 11, angle = 0, vjust = 1, hjust = 0.5),
        axis.text.y = element_text(color="black", size = 11),
        legend.position="none")
  
plot14 
```


```{r}
#Testing whether a platform's time in the tank affected whether prey was consumed  
prey = read_excel("Eb_PA_Consumption.xlsx")
summary(prey)

##subsetting 2020 fish 
prey20 <- subset(prey, prey$YEAR == "2020")

##Is all platform's time in tank a significant predictor of total live prey consumed? 
shapiro.test(prey20$TOTAL_LIVE)
kruskal.test(TOTAL_LIVE ~ ALL_PLATFORMS_TANKTIME, data = prey20) ##p-value = 0.3044

a11<- aov(data = prey20, TOTAL_LIVE ~ as.factor(ALL_PLATFORMS_TANKTIME) * as.factor(MODALITY))
summary(a11) #NS


##Is all platform's time in tank a significant predictor of total dead prey consumed? 
shapiro.test(prey20$TOTAL_DEAD)
kruskal.test(TOTAL_DEAD ~ ALL_PLATFORMS_TANKTIME, data = prey20) ##p-value = 0.8085

a12<- aov(data = prey20, TOTAL_DEAD ~ as.factor(ALL_PLATFORMS_TANKTIME) * as.factor(MODALITY))
summary(a12) #NS



##Individual platform analyis for does platform time in tank affeect whether live or dead prey was consumed? 

#Live prey 
kruskal.test(P1_LIVE ~ P1_TANKTIME, data = prey20) #NS, p-value = 0.2363

a3<- aov(data = prey20, P1_LIVE ~ P1_TANKTIME * MODALITY)
summary(a3) #NS

kruskal.test(P2_LIVE ~ P2_TANKTIME, data = prey20) #NS, p-value = 0.1017

a4<- aov(data = prey20, P2_LIVE ~ P2_TANKTIME * MODALITY)
summary(a4) #NS


kruskal.test(P3_LIVE ~ P3_TANKTIME, data = prey20) ##NS, p-value = 0.2282
a5<- aov(data = prey20, P3_LIVE ~ P3_TANKTIME * MODALITY)
summary(a5) #NS

kruskal.test(P4_LIVE ~ P4_TANKTIME, data = prey20) #NS, p-value = 0.2399
a6<- aov(data = prey20, P4_LIVE ~ P4_TANKTIME * MODALITY)
summary(a6) #NS 

#Dead prey 
kruskal.test(P1_DEAD ~ P1_TANKTIME, data = prey20) #NS, p-value = 0.2739
a7<- aov(data = prey20, P1_DEAD ~ as.factor(P1_TANKTIME) * as.factor(MODALITY))
summary(a7) #Modality & interaction significant 
ph3 <- glht(a7, linfct = mcp("as.factor(MODALITY)" = "Tukey"))
summary(ph3) #dead prey consumed varied significantly in dark trials vs. dark trials w/ LL ablated, light trials w/ LL ablated vs. dark trials w/ LL ablated, and dark trials w/ LL ablated vs. light trials w/ LL 

ph6 <- glht(a7, linfct = mcp("as.factor(P1_TANKTIME)" = "Tukey"))
summary(ph6) #NS 

kruskal.test(P2_DEAD ~ P2_TANKTIME, data = prey20) #NS, p-value = 0.7295
a8<- aov(data = prey20, P2_DEAD ~ P2_TANKTIME * MODALITY) 
summary(a8) #NS


kruskal.test(P3_DEAD ~ P3_TANKTIME, data = prey20) #NS, p-value = 0.6296
a9<- aov(data = prey20, P3_DEAD ~ P3_TANKTIME * MODALITY)
summary(a8) #NS

kruskal.test(P4_DEAD ~ P4_TANKTIME, data = prey20) #NS, p-value = 0.1939
a10<- aov(data = prey20, P4_DEAD ~ P4_TANKTIME * MODALITY)
summary(a10) #NS
```

